Understanding Factors Contributing to Application Approval in Urban Planning Decisions (2016-2024)

### 2024-12-02 paper/paper.qmd
Upload abstract and introduction

---
title: "Understanding Factors Contributing to Application Approval in Urban Planning Decisions (2016-2024)"
subtitle: "Analysis using data from Toronto’s Committee of Adjustment applications using Bayesian Logistic Regression"
author: Angel Xu
thanks: "Code and data are available at: [https://github.com/Anjojoo/Committee-of-Adjustment-Applications](https://github.com/Anjojoo/Committee-of-Adjustment-Applications)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

@@

@@

## Measurement

@@

## Data Cleaning

@@

## Features

@@

2. Confusion Metrix

3. Sensitivity Analysis
The sensitivity of the model was assessed by varying the priors for the coefficients to examine whether the results were robust to different prior specifications. The predictions remained stable across different prior choices, indicating that the model is not overly influenced by prior assumptions.

4. Predictive Power (AUC-ROC)
The Area Under the Receiver Operating Characteristic Curve (AUC-ROC) was computed to evaluate the model's ability to discriminate between approved and refused applications. A high AUC-ROC value (>0.8) confirmed the model's strong discriminatory performance.

# Results

@@

# Discussion

>>>>

---
title: "Understanding Factors Contributing to Application Approval in Urban Planning Decisions (2016-2024)"
subtitle: "Analysis using data from Toronto’s Committee of Adjustment applications using Bayesian Logistic Regression"
author: Angel Xu
thanks: "Code and data are available at: [https://github.com/Anjojoo/Committee-of-Adjustment-Applications](https://github.com/Anjojoo/Committee-of-Adjustment-Applications)."
date: today
date-format: long
abstract: "This paper analyzes urban planning application approvals submitted to Toronto's Committee of Adjustment from 2016 to 2024, focusing on key predictors such as application type, submission year, and planning district using Bayesian logistic regression. The analysis shows that consents are more likely to be approved than minor variances, with approval rates generally increasing over time, particularly from 2017 onward, though fluctuations are observed in certain years. Geographic disparities were identified, with North York showing the highest likelihood of approval, while Scarborough faced stricter reviews. These findings enhance understanding of urban planning decision-making processes, offering insights into temporal trends, geographic variability, and areas for equitable policy adjustments."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

Urban planning decisions shape the development of cities, influencing how communities grow and adapt to changing social, economic, and environmental conditions. In Toronto, one of Canada’s largest and most diverse cities, the approval of planning applications directly impacts growth, housing availability, and community design. Despite the importance of this process, limited research has explored the specific factors that influence approval decisions, particularly for minor variances and consents. This lack of understanding presents challenges for creating equitable and effective policies that balance the needs of stakeholders while adhering to regulatory frameworks.

This paper examines the factors influencing approval decisions for planning applications submitted to the Committee of Adjustment in Toronto between 2016 and 2024. By analyzing application type, year of submission, and planning district, this study identifies patterns and trends in approval rates. Using a statistical modeling approach, it investigates how administrative processes, policy changes, and regional differences affect these outcomes over time. This analysis addresses a gap in urban planning research, providing a systematic evaluation of the approval process and its determinants.

The findings indicate that consents are more likely to be approved than minor variances, reflecting differences in the scrutiny applied to these application types. Approval rates show notable temporal trends, with the highest approval likelihood observed in 2018, where odds of approval were 161% higher compared to the baseline year of 2016. Significant increases are also evident in 2017 and 2020, with approval rates showing consistent improvement over time, despite some fluctuations. Additionally, geographic disparities are evident, with North York exhibiting the highest approval probabilities, while Scarborough demonstrates lower likelihoods, potentially due to district-specific planning challenges. The difference in approval rates across Toronto’s planning districts suggests localized factors influence decision-making. These results enhance understanding of the approval process and highlight areas where decision-making could become more consistent and equitable.

The remainder of this paper is structured as follows. @sec-data introduces the overview of the data (@sec-data-overview), measurement (@sec-data-measurement), data cleaning process (@sec-data-cleaning), data features (@sec-data-feature), as well as explanations, descriptions, table and graph summaries of outcome (@sec-data-outcome) and predictor variables (@sec-data-predictor) of the study. @sec-model explains the modeling process in detail, including procedure of model set-up, model justifications, model assumptions, and potential limitations. Then, @sec-result highlights the prediction outcome and results by model summary and plots. Finally, @sec-dis includes discussions based on the results, linking it to the real world and also states some strengths and weakness of the study.

@@

## Overview {#sec-data-overview}

@@

## Measurement {#sec-data-measurement}

@@

## Data Cleaning {#sec-data-cleaning}

@@

## Features {#sec-data-feature}

@@

2. **Confusion Metrix**
The confusion matrix was computed to evaluate the model's classification performance, providing measures such as accuracy, sensitivity, and specificity. This validation assesses the model's predictive capabilities and its performance in correctly classifying approval decisions.

# Results{#sec-result}

@@

# Discussion{#sec-dis}
